Actuator line modeling of vertical-axis turbines

### 2018-05-06 paper.tex
Finish addressing reviewer 4

After the force on the ALE from the flow is calculated, it is then projected
back onto the flow field as a source term in the momentum equation. To avoid
instability due to steep gradients, the source term is tapered from its maximum
value away from the element location by means of a spherical Gaussian function.
The width of this function $\eta$ is controlled by a single parameter
$\epsilon$, which is then multiplied by the actuator line element force and
imparted on a cell with distance $| \vec{r} |$ from the actuator line element
quarter chord location:
\begin{equation}
    \eta = \frac{1}{\epsilon^3 \pi^{3/2}} \exp
    \left[ - \left( \frac{| \vec{r} |}{\epsilon} \right)^2 \right].
    \label{eq:projection}
\end{equation}

@@

In the context of a turbine---especially a cross-flow turbine---the actuator
lines will encounter unsteady conditions, both in their angle of attack and
relative velocity. These conditions necessitate the use of unsteady aerodynamic
models to augment the static foil characteristics, both to capture the time
resolved response of the attached flow loading and effects of flow acceleration,
also know as added mass. Furthermore, the angles of attack encountered by a CFT
blade will often be high enough to encounter dynamic stall (DS). It is therefore
necessary to model both unsteady attached and detached flow to obtain accurate
loading predictions.

@@

Helmholtz's second vortex theorem states that vortex lines may not end in a
fluid, but must either form closed loops or extend to boundaries. Consequently
the lift distribution due to the bound vortex from foils of finite span must
drop to zero at the tips. One popular end effects correction was developed by
Glauert~\cite{Glauert1935} for the blade element analysis of axial-flow rotors.
However this correction depends on rotor parameters---tip speed ratio, number of
blades, element radius, tip flow angle---that do no necessarily translate
directly to the geometry and flow environment of a cross-flow rotor. Therefore,
a more general end effects model was sought.

@@

    \caption{Weighted average momentum recovery terms for the actuator disk (AD)
        simulation from \cite{Bachant2015-JoT}, UNH-RVAT actuator line model with a
        $k$--$\epsilon$ RANS closure, the two 3-D blade resolved RANS models
        described in \cite{Bachant2016-BR-CFD}, and the experiments reported in
        \cite{Bachant2015-JoT}.}

>>>>

After the force on the actuator line element from the flow is calculated,
it is then projected
back onto the flow field as a source term in the momentum equation. To avoid
instability due to steep gradients, the source term is tapered from its maximum
value away from the element location by means of a spherical Gaussian function.
The width of this function $\eta$ is controlled by a single parameter
$\epsilon$, which is then multiplied by the actuator line element force and
imparted on a cell with distance $| \vec{r} |$ from the actuator line element
quarter chord location:
\begin{equation}
    \eta = \frac{1}{\epsilon^3 \pi^{3/2}} \exp
    \left[ - \left( \frac{| \vec{r} |}{\epsilon} \right)^2 \right].
    \label{eq:projection}
\end{equation}

@@

In the context of a turbine---especially a cross-flow turbine---the actuator
lines will encounter unsteady conditions, both in their angle of attack and
relative velocity. These conditions necessitate the use of unsteady aerodynamic
models to augment the static foil characteristics, both to capture the time
resolved response of the attached flow loading and effects of flow acceleration,
also known as added mass. Furthermore, the angles of attack encountered by a CFT
blade will often be high enough to encounter dynamic stall (DS). It is therefore
necessary to model both unsteady attached and detached flow to obtain accurate
loading predictions.

@@

Helmholtz's second vortex theorem states that vortex lines may not end in a
fluid, but must either form closed loops or extend to boundaries. Consequently
the lift distribution due to the bound vortex from foils of finite span must
drop to zero at the tips. One popular end effects correction was developed by
Glauert~\cite{Glauert1935} for the blade element analysis of axial-flow rotors.
However this correction depends on rotor parameters---tip speed ratio, number of
blades, element radius, tip flow angle---that do not necessarily translate
directly to the geometry and flow environment of a cross-flow rotor. Therefore,
a more general end effects model was sought.

@@

    \caption{
        Weighted average momentum recovery terms for the actuator disk (AD)
        simulation from \cite{Bachant2015-JoT}, UNH-RVAT actuator line model
        with a $k$--$\epsilon$ RANS closure, the two 3-D blade resolved RANS
        models described in \cite{Bachant2016-BR-CFD} ($k$--$\omega$ SST and
        Spalart--Allmaras, SA), and the experiments reported in
        \cite{Bachant2015-JoT}.
    }
